---
title: "02_cogtask_cleaning"
author: "Levente Rónai"
output: html_document
---

# loading packages

```{r, include=F, warning=F, message=F}

library(corrplot)
library(readxl)
library(xts)
library(psych)
library(qgraph)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(car)
library(lubridate)
library(knitr)
library(git2r)
library(psycho)

```

# reading datasets

```{r, warning=F, message=F}

options("scipen" = 100, "digits" = 4)

dat_nback   <- read.csv("raw_data/merged_file_nback.csv")

dat_gonogo   <- read.csv("raw_data/merged_file_gonogo.csv")

```

# selection and renaming of relevant variables 

(session id, type of stimuli [go/no-go], reaction time, type of cogtask, date of trial created, number of stimulus within one trial)

```{r}


dat_nback <- 
  dat_nback %>% 
  as_tibble() %>% 
  select(session_esm = code,
          trial_type = trialType,
                  rt = nBackClick.time,
                task = expName,
        date_cogtask = date,
             n_trial = trials.thisN)


dat_gonogo <- 
  dat_gonogo %>% 
  as_tibble() %>% 
  select(session_esm = code,
          trial_type = trialType,
                  rt = nBackClick.time,
                task = expName,
        date_cogtask = date,
             n_trial = trials.thisN)


# fit the two datasets together

dat_cogtask <- 
  dat_gonogo %>% 
  rbind(dat_nback)

```

```{r}
dat_cogtask %>% 
  group_by(session_esm, date_cogtask, task) %>% 
  count() %>% 
  group_by(task) %>% 
  summarise(
     mean_number_of_trials = mean(n), 
     sd_of_number_of_trials = sd(n)
  )
```

# cleaning

```{r}
dat_cogtask <- 
  
  dat_cogtask %>% 
  # clean the number of the stimuli (the range is between 0-59, every 60th is an NA row)
  mutate(n_trial = n_trial + 1) %>% 
  # (1-60, every 61st is an NA row)
  filter(!is.na(n_trial)) %>% 
  # clean rt - empty cells, empty brackets, or real rt values in brackets   
  mutate(rt = ifelse(rt == "", NA_real_, as.character(rt)),
         rt = ifelse(rt == "[]", NA_real_, as.character(rt)),
         rt = str_sub(rt, 2, -2))

```

```{r}

dat_cogtask %>% 
  group_by(session_esm, date_cogtask, task) %>% 
  count() %>% 
  group_by(task) %>% 
  summarise(
     mean_number_of_trials = mean(n), 
     sd_of_number_of_trials = sd(n)
  )
```

# import and transform of test trial

```{r, warning=F, message=F}
# there are reaction times with multiple values in one cell

dat_cogtask %>% 
  arrange(rt) # %>% view()

```


```{r, warning=F, message=F}

# test-trial made with two techniques of responding:  

dat_test   <- read.csv("raw_data/test_trial.csv")

  # in the first part, I have performed frequent "knocking" with my finger on the screen without stimulus-processing
  # in the second part, I have held my finger permanently on the screen without raising it

```


```{r, message=F}

dat_test <-
  dat_test %>%
  as_tibble() %>%
  select(session_esm = code,
          trial_type = trialType,
                  rt = nBackClick.time,
                task = expName,
        date_cogtask = date,
             n_trial = trials.thisN) # %>% view()


# it seems obvious that bugged values in rts are due to random clicking/taping

```

# separate multiple rt values into multiple columns 

```{r, message = F, warning = F}

dat_cogtask <- 
  dat_cogtask %>% 
  
  separate(rt, c("rt_1st_part", "rt_2nd_part", "rt_3rd_part"), sep = ",") %>% 
  mutate_at(vars(rt_1st_part:rt_3rd_part),
            .funs = list(num = ~ round(as.numeric(.), digits = 3))) %>%
  select(session_esm, trial_type, contains("rt"), everything(), -rt_1st_part, -rt_2nd_part, -rt_3rd_part)

```

# check whether bugged responses do form a sequence and belong together or just random values   

```{r}

dat_cogtask %>% 
  mutate(
    test = 
      ifelse(is.na(rt_3rd_part_num),
             rt_1st_part_num < rt_2nd_part_num,
             (rt_1st_part_num < rt_2nd_part_num) & (rt_2nd_part_num < rt_3rd_part_num))) %>%
  select(test) %>% 
  table()

```

# screen delayed, but not fake responses

```{r}

dat_cogtask <- 
  
  dat_cogtask %>% 
  
  # create lagged variables to identify late responses
  
  group_by(session_esm) %>% 
  mutate(lagged_trial_type = lag(trial_type),
         lagged_rt_1st_part_num = lag(rt_1st_part_num)) %>% 
  ungroup() %>%

  # screen for late responses 
  
  mutate(late = ifelse(
                  lagged_trial_type == 1 &        # if the previous stim is go
                  is.na(lagged_rt_1st_part_num) & # but the previous rt is missing  
                  !is.na(rt_2nd_part_num) &       # if there is a second value of rt
                  is.na(rt_3rd_part_num) &        # but there is no third value of rt    
                  rt_1st_part_num < 0.2,          # if it is way too fast, it can be regarded as a late response
                1,
                0)
         ) %>% 
  
  # subtract late responses from the second value of rt to get valid response times 
    
  mutate(rt_1st_part_num = ifelse(late == 1,
                                  as.numeric(rt_2nd_part_num - rt_1st_part_num),
                                  rt_1st_part_num),
         rt_2nd_part_num = ifelse(late == 1,
                                  NA,
                                  rt_2nd_part_num)) %>% 
  
  # mark the remaining multiplied values as bugged

  mutate(bugged = ifelse(!is.na(rt_2nd_part_num),
                         TRUE,
                         FALSE)) %>%
  
  select(-lagged_trial_type, -lagged_rt_1st_part_num, - late)

```

# rate of bugged trials per participant


```{r, warning=F}

# identify trials with higher than 50% bugged/normal response rate

fake_session <-
  dat_cogtask %>% 
  filter(!is.na(rt_1st_part_num)) %>% # consider only responded trials
  group_by(session_esm, date_cogtask) %>% # group by participant and session 
  summarise(
    n_bugged    = sum(bugged),
    n_normal    = sum(!bugged),
    rate        = n_bugged / n_normal, # ratio of bugged to normal responses 
  ) %>% 
  arrange(desc(n_bugged)) %>%


  mutate(fake_session = ifelse(rate > 0.5, 
                            1,
                            0)) %>% 
  ungroup() %>% 
  select(session_esm, date_cogtask, fake_session, rate)


```

```{r}

dat_cogtask <-
  dat_cogtask %>%
  left_join(fake_session, by = c("session_esm", "date_cogtask"))

```

# screening in 3 + 1 steps

```{r}

# Three steps

dat_cogtask <- 
  dat_cogtask %>% 
         
  filter(
    # 1st step: possible sessions that are completely fake (more than 50% bugged/normal response rate): 
    fake_session == 0 &  
      
      # 2nd step: bugged (NOT delayed!) responses (two or more reaction time values for one stimuli)
      bugged == FALSE &
      
      # 3rd step: unlikely quick responses
      rt_1st_part_num > 0.2 | 
      
      # keep trials with no responses from non-fake sessions
      (is.na(rt_1st_part_num) & fake_session == 0)) %>% 
  
  
select(session_esm, task, trial_type, rt = rt_1st_part_num, date_cogtask, n_trial, fake_session, bugged)


```

```{r}

# check the number of trials within sessions

dat_cogtask %>% 
  group_by(session_esm, date_cogtask, task) %>% 
  count() %>% 
  group_by(task) %>% 
  summarise(
    min_n_of_trials   = min(n),
    max_n_of_trials   = max(n),
    mean_n_of_trials  = mean(n), 
    sd_of_n_of_trials = sd(n),
    sum_n_of_trials   = sum(n)

  )

```


```{r}

# Extra screening step - filter trials without a response given (by this we also filter way too faked sessions, see summary below)   

dat_cogtask <- 
  dat_cogtask %>% 
  group_by(session_esm, date_cogtask) %>% 
  mutate(count_of_trial = n_distinct(n_trial),
         mean_rt_trial = mean(rt, na.rm = T),
         to_delete = ifelse(is.na(mean_rt_trial),
                            1,
                            0)
         ) %>% 
  ungroup()

dat_cogtask$to_delete %>% table()

# 5.6 % of the sample will be deleted

```

```{r}

dat_cogtask <- 
  dat_cogtask %>% 
  filter(to_delete == 0) %>% 
  select(-mean_rt_trial, -to_delete)

```


```{r}

dat_cogtask %>% 
  group_by(session_esm, date_cogtask, task) %>% 
  count() %>% 
  group_by(task) %>% 
  summarise(
    min_n_of_trials   = min(n),
    max_n_of_trials   = max(n),
    mean_n_of_trials  = mean(n), 
    sd_of_n_of_trials = sd(n),
    sum_n_of_trials   = sum(n)
  )

```


# calculate and join of dprime scores

```{r}

# keep raw data before the computation of dprime scores

write_csv(dat_cogtask, "processed_data/dat_cogtask_raw.csv")

```


```{r}

dat_cogtask <-
  dat_cogtask %>%
  mutate(hit     = ifelse(trial_type == 1 & !is.na(rt),  # hits
                      1,
                      0),
         fa      = ifelse(trial_type == 0 & !is.na(rt),  # false alarms
                      1,
                      0),
         miss    = ifelse(trial_type == 1 & is.na(rt),   # misses
                      1,
                      0),
         cr      = ifelse(trial_type == 0 & is.na(rt),   # correct rejections
                      1,
                      0),
         corr_rt = ifelse(trial_type == 1,
                      rt,
                      NA)
  ) %>%
  
  group_by(session_esm, date_cogtask) %>% 
  summarise(
    n_of_trials    = mean(count_of_trial),
    task           = max(as.character(task)),
    
    corr_rt_median = median(corr_rt, na.rm = T),
    
    n_hit          = sum(hit),
    n_fa           = sum(fa),
    n_miss         = sum(miss),
    n_cr           = sum(cr),
    n_targets      = n_hit + n_miss,
    n_distractors  = n_fa + n_cr)


```

```{r}

dprime <- 
  psycho::dprime(n_hit         = dat_cogtask$n_hit,
                 n_fa          = dat_cogtask$n_fa,
                 n_miss        = dat_cogtask$n_miss,
                 n_cr          = dat_cogtask$n_cr,
                 n_targets     = dat_cogtask$n_targets,
                 n_distractors = dat_cogtask$n_distractors)

```


```{r}

dat_cogtask <-
  dat_cogtask %>%
  cbind(dprime) %>% 
  select(everything(), -c(n_hit:n_distractors), -c(aprime:c)) 

```

# convert timestamps to beeps

```{r}

dat_cogtask <- 
  dat_cogtask %>% 
  # keep original timestamp - will be used as a key to calculate within-session reliability
  mutate(date_cogtask_orig = date_cogtask) %>% 

  # convert the date/time of tasks to dates and beeps
  
  # split date_cogtask to date and hours/mins
  separate(date_cogtask, c("date_cogtask", "time_cogtask"), sep = "_") %>%
  group_by(time_cogtask) %>% 
  mutate(time_cogtask = unlist(strsplit(time_cogtask, split = ".", fixed = T))[1],
         time_cogtask = lubridate::hm(time_cogtask),
         date_cogtask = as.Date(date_cogtask)) %>% 
  ungroup() %>% 

  # create beeps
  
  mutate(
    
    # initialize column with double (aka real number) NA values
    beep         = NA_real_, 
    
    # convert time to minutes passed since midnight
    hour         = lubridate::hour(time_cogtask),
    minutes      = lubridate::minute(time_cogtask),
    cog_mins     = (hour * 60 + minutes)
  ) %>% 
  
  # calculate beep based on timestamp of the creation of the cogtask trial
  mutate(
    beep =     
      case_when(
        between(cog_mins, 420, 569)   ~ 1,    # beep around 08:00
        between(cog_mins, 570, 689)   ~ 2,    # beep around 10:00
        between(cog_mins, 690, 809)   ~ 3,    # beep around 12:00
        between(cog_mins, 810, 929)   ~ 4,    # beep around 14:00
        between(cog_mins, 930, 1049)  ~ 5,    # beep around 16:00
        between(cog_mins, 1050, 1169) ~ 6,    # beep around 18:00
        between(cog_mins, 1170, 1289) ~ 7,    # beep around 20:00
        between(cog_mins, 1290, 1409) ~ 8)    # beep around 22:00
  ) %>% 
  arrange(session_esm)  %>% 
  select(-cog_mins, -time_cogtask) %>%
  distinct()

```


```{r}

# there are some duplicated results within the same beep

dat_cogtask %>%
  group_by(session_esm, date_cogtask, task) %>%
  mutate(dupli = ifelse(duplicated(beep) | duplicated(beep, fromLast = TRUE),
                        1,
                        NA)) %>%
  filter(dupli == 1) # %>% view()

# we keep the first results only from the duplicated ones

dat_cogtask <- 
  dat_cogtask %>% 
  group_by(session_esm, date_cogtask, task) %>%
  filter(!duplicated(beep))

```

# spread data regarding type of cogtask

```{r, warning=F}

dat_cogtask <-
  dat_cogtask %>%
  mutate(task = ifelse(task == "go-nogo",
                       "gonogo",
                       "nback")) %>% 
  

  nest(dprime, beta, corr_rt_median, minutes, .key = 'value_col') %>%
  spread(key = task, value = value_col)  %>% 
  unnest(gonogo, nback, .sep = '_')


```
# define days

```{r}

big_data <- read.csv("processed_data/big_data.csv")

start_dates <- 
  big_data %>% 
  select(session_esm, start_date) %>% 
  distinct()

dat_cogtask <- 
  dat_cogtask %>% 
  left_join(start_dates, by = c("session_esm")) %>% 
  group_by(session_esm) %>% 
  mutate(day = as.numeric(date_cogtask - as.Date(start_date))) %>% 
 # select(-start_date) %>% 
  ungroup()

```

# further screening

```{r}

# filter sessions created before the offical start of the study (test sessions)

dat_cogtask <- 
  dat_cogtask %>% 
  filter(date_cogtask > "2021-04-11")

```

# complete days and beeps per participants

```{r}

dat_cogtask <- 
  dat_cogtask %>%
  
  # define sets of possible values for variables to complete 
  mutate(
    day  = factor(day, levels = 1:28),
    beep = factor(beep, levels = 1:8)
  ) %>% 
  group_by(session_esm) %>% 
  tidyr::complete(day, beep)

```

# fill nback and gonogo tasks into the same beep that are belonging together 

```{r}
dat_cogtask <- 
  dat_cogtask %>% 
  select(session_esm:date_cogtask, date_cogtask_orig, hour, gonogo_dprime, gonogo_corr_rt_median, gonogo_beta,
         nback_dprime, nback_corr_rt_median, nback_beta, gonogo_minutes, nback_minutes) %>%
  filter(!is.na(day)) %>% 
  group_by(session_esm, day, beep) %>%
  fill(c(gonogo_dprime:nback_beta, gonogo_minutes, nback_minutes), .direction = "updown") %>%
  distinct() %>%
  ungroup()

```


```{r}
dat_cogtask <- 
  dat_cogtask %>%   
  # screen two beeps in which the two cogtasks have not occurred in the same hour due to an error in pavlovia/formr.
  # this results the exclusion of olny two cases, no systematic screening is needed
  
  group_by(session_esm, day) %>%
  mutate(filter_1 = ifelse(duplicated(beep) # 
                           & session_esm == "ZXy6-W3UxKqK95cydAuqjBsgBVwC7vmWuKNH56mDPBmwEST1ypn4IgjuDubaNumm",
                           1,
                           0),
         filter_2 = ifelse(duplicated(beep, fromLast = TRUE) 
                           & session_esm == "c3QV9DlxJpiAScLV2XXT729aBSTXbcP16lBVL9vJgbmwTgteduWI6EpekmuhyGlP",
                           1,
                           0),
         to_discard = ifelse(((filter_2 == 0 & filter_1 == 1)
                              | (filter_2 == 1 & filter_1 == 0)),
                             1,
                             0
                             )) %>% 
  filter(to_discard == 0) %>% 
  ungroup() %>% 
  select(-filter_1, -filter_2, -to_discard)

```

# write data

```{r}

write_csv(dat_cogtask,"processed_data/dat_cogtask.csv")

```

