---
title: "04_descriptive_stats"
author: "Flóra Hann"
output: html_document
---

```{r, message = F, warning = F}

options(scipen = 999)

### loading packages

library(psych)
library(tidyverse)
library(dplyr)
library(car)
library(data.table)
library(lubridate)
library(git2r)
library(naniar)
library(hms)
library(nortest)

```


# importing and joining datasets of self-reports and behavioural assessments

```{r}

### import datasets of self-reports and behavioural assessments

big_data <- read.csv("processed_data/big_data.csv")

dat_pre_cleaned <- read.csv("raw_data/dat_pre_cleaned.csv")

dat_cogtask <- read.csv("processed_data/dat_cogtask.csv")

dat_esm_cleaned_unfiltered <- read.csv("raw_data/dat_esm_cleaned_unfiltered.csv")

model_dat <- read.csv("processed_data/model_dat.csv")

dat_cleaned <- read.csv("processed_data/dat_cleaned.csv")

big_data_screened <- read.csv("processed_data/big_data_screened.csv")

model_dat_filtered <- read.csv("processed_data/model_dat_filtered.csv")

```

# sample demographics

## cross-sectional sample

```{r}

# number and age of participants  
  
  dat_pre_cleaned %>% 
    filter(!is.na(session)) %>% 
    select(session, age) %>% 
    distinct() %>% 
    summarise(N         = n(),
              Mean      = round(mean(age), digits = 2),
              Median    = round(median(age), digits = 2),
              SD        = round(sd(age), digits = 2),
              Range_min = min(age),
              Range_max = max(age))


# gender

  # 1 = Female
  # 2 = Male

  # NA:
  # 3 = Other

dat_pre_cleaned %>% 
  filter(!is.na(session)) %>% 
  select(session, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table()

dat_pre_cleaned %>% 
  filter(!is.na(session)) %>% 
  select(session, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)


# education

  # 1 = Primary school or lower
  # 3 = Vocational school without high school diploma
  # 4 = High school diploma or equivalent 
  # 5 = Bachelor's or Master's degree
  # 6 = Doctorate (PhD)
  # 7 = Other
  
dat_pre_cleaned %>% 
  filter(!is.na(session)) %>% 
  select(session, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table()

dat_pre_cleaned %>% 
  filter(!is.na(session)) %>% 
  select(session, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

```

## proceeded to the ESM phase

```{r}

# number of participants

dat_esm_cleaned_unfiltered %>%
    filter(!is.na(session_esm), (!is.na(ended_esm))) %>%
    select(session_esm) %>%
    distinct() %>%
    summarise(N = n_distinct(session_esm))

```

## not included

```{r}

# identify participants not included in the model

model_dat_participants <- model_dat_filtered %>% 
  select(id_esm:start_date) %>% 
  mutate(
    gender =
      if_else(
        gender == "Female",
        1,
        2
      )
    ) %>%
  distinct()

dat_not_included <- dat_pre_cleaned %>%
  filter(!is.na(session)) %>% 
  anti_join(model_dat_participants, by = c("session" = "session_esm"))


# number and age of participants  

dat_not_included %>% 
  select(session, age) %>% 
  distinct() %>% 
  summarise(N         = n_distinct(session),
            Mean      = round(mean(age), digits = 2),
            Median    = round(median(age), digits = 2),
            SD        = round(sd(age), digits = 2),
            Range_min = min(age),
            Range_max = max(age))

# gender

  # 1 = Female
  # 2 = Male

  # NA:
  # 3 = Other

dat_not_included %>%
  select(session, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table()

dat_not_included %>%
  select(session, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

# education

  # 1 = Primary school or lower
  # 3 = Vocational school without high school diploma
  # 4 = High school diploma or equivalent 
  # 5 = Bachelor's or Master's degree
  # 6 = Doctorate (PhD)
  # 7 = Other
  
dat_not_included %>% 
  select(session, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table()

dat_not_included %>% 
  select(session, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

```

## final sample

```{r}

# number and age of participants  

model_dat_filtered %>% 
  filter(!is.na(id_esm)) %>% 
  select(id_esm, age) %>% 
  distinct() %>% 
  summarise(N         = n_distinct(id_esm),
            Mean      = round(mean(age), digits = 2),
            Median    = round(median(age), digits = 2),
            SD        = round(sd(age), digits = 2),
            Range_min = min(age),
            Range_max = max(age))

# gender

  # 1 = Female
  # 2 = Male

  # NA:
  # 3 = Other

model_dat_filtered %>% 
  filter(!is.na(id_esm)) %>% 
  select(id_esm, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table()

model_dat_filtered %>% 
  filter(!is.na(id_esm)) %>% 
  select(id_esm, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

# education

  # 1 = Primary school or lower
  # 3 = Vocational school without high school diploma
  # 4 = High school diploma or equivalent 
  # 5 = Bachelor's or Master's degree
  # 6 = Doctorate (PhD)
  # 7 = Other
  
model_dat_filtered %>% 
  filter(!is.na(id_esm)) %>% 
  select(id_esm, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table()

model_dat_filtered %>% 
  filter(!is.na(id_esm)) %>% 
  select(id_esm, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

```

# descriptives

## descriptives of ESM surveys in final sample

```{r}

# total number of completed esm surveys

model_dat_filtered %>%
  distinct() %>% 
  summarise(n_id  = length(unique(id_esm)),
            n_obs = length(event_unpleasantness_centered))

#...and per capita

big_data_screened %>%
  distinct() %>% 
  filter(!is.na(date_esm)) %>% 
  group_by(id_esm) %>% 
  summarise(n_obs = length(date_esm)) %>% 
  ungroup() %>% 
  summarise(max_n_obs    = max(n_obs),
            min_n_obs    = min(n_obs),
            mean_n_obs   = mean(n_obs),
            median_n_obs = median(n_obs),
            sd_n_obs     = sd(n_obs))



  model_dat_ESM <- model_dat %>% 
  filter(!is.na(event_pleasantness))
  
  observations_not_included_ESM <- model_dat_ESM %>% 
  anti_join(model_dat_filtered, by = c("id_esm", "day", "beep"))


# descriptives of event pleasantness

  # in observations included
  
  model_dat_filtered <- model_dat_filtered %>%
    mutate(event_unpleasantness = 8-event_pleasantness)
  
  model_dat_filtered %>% 
    select(id_esm, day, beep, event_unpleasantness) %>% 
    summarise(
      mean_evunpleasant   = mean(event_unpleasantness),
      median_evunpleasant = median(event_unpleasantness),
      sd_evunpleasant     = sd(event_unpleasantness),
      min_evunpleasant    = min(event_unpleasantness),
      max_evunpleasant    = max(event_unpleasantness)
    )
  
  
  # in observations not included
    
  observations_not_included_ESM <- observations_not_included_ESM %>% 
    mutate(event_unpleasantness = 8-event_pleasantness)

  observations_not_included_ESM %>% 
  select(id_esm, day, beep, event_unpleasantness) %>% 
    summarise(
      mean_evunpleasant   = mean(event_unpleasantness),
      median_evunpleasant = median(event_unpleasantness),
      sd_evunpleasant     = sd(event_unpleasantness),
      min_evunpleasant    = min(event_unpleasantness),
      max_evunpleasant    = max(event_unpleasantness)
    )

```

## descriptives of cogtasks in final sample

### go/no-go

```{r}

# total number of completed go/no-go tasks

big_data_screened %>%
  distinct() %>%
  filter(!is.na(gonogo_dprime)) %>% 
  summarise(n_id  = length(unique(id_esm)),
            n_obs = length(gonogo_dprime))

# ...and per capita

big_data_screened %>%
  distinct() %>% 
  filter(!is.na(gonogo_dprime)) %>% 
  group_by(id_esm) %>% 
  summarise(n_obs = length(id_esm)) %>%
  ungroup() %>% 
  summarise(max_n_obs    = max(n_obs),
            min_n_obs    = min(n_obs),
            mean_n_obs   = mean(n_obs),
            median_n_obs = median(n_obs),
            sd_n_obs     = sd(n_obs))

# descriptives of go/no-go performance

model_dat_filtered %>%
  select(id_esm, day, beep, gonogo_dprime) %>%
  summarise(
    mean_gonogo_dprime   = mean(gonogo_dprime),
    median_gonogo_dprime = median(gonogo_dprime),
    sd_gonogo_dprime     = sd(gonogo_dprime),
    min_gonogo_dprime    = min(gonogo_dprime),
    max_gonogo_dprime    = max(gonogo_dprime)
  )

# descriptives of go/no-go performance in observations not included

  # identifying observations not included

  model_dat_gonogo <- model_dat %>% 
    filter(!is.na(gonogo_dprime))
  
  observations_not_included_gonogo <- model_dat_gonogo %>% 
    anti_join(model_dat_filtered, by = c("id_esm", "day", "beep"))
  
  observations_not_included_gonogo %>%
  select(id_esm, day, beep, gonogo_dprime) %>%
  summarise(
    mean_gonogo_dprime   = mean(gonogo_dprime),
    median_gonogo_dprime = median(gonogo_dprime),
    sd_gonogo_dprime     = sd(gonogo_dprime),
    min_gonogo_dprime    = min(gonogo_dprime),
    max_gonogok_dprime    = max(gonogo_dprime)
  )

```


### n-back

```{r}
   
# total number of completed n-back tasks
   
big_data_screened %>%
  distinct() %>% 
  filter(!is.na(nback_dprime)) %>% 
  summarise(n_id = length(unique(id_esm)),
            n_obs = length(nback_dprime))
   
# ...and per capita
  
  big_data_screened %>%
    distinct() %>% 
    filter(!is.na(nback_dprime)) %>% 
    group_by(id_esm) %>% 
    summarise(n_obs = length(id_esm)) %>% 
    ungroup() %>% 
    summarise(max_n_obs = max(n_obs),
              min_n_obs = min(n_obs),
              mean_n_obs = mean(n_obs),
              median_n_obs = median(n_obs),
              sd_n_obs = sd(n_obs))
  
# descriptives of n-back performance in observations included

model_dat_filtered %>%
  select(id_esm, day, beep, nback_dprime) %>%
  summarise(
    mean_nback_dprime   = mean(nback_dprime),
    median_nback_dprime = median(nback_dprime),
    sd_nback_dprime     = sd(nback_dprime),
    min_nback_dprime    = min(nback_dprime),
    max_nback_dprime    = max(nback_dprime)
  )

# descriptives of n-back performance in observations not included

  # identifying observations not included

  model_dat_nback <- model_dat %>% 
    filter(!is.na(nback_dprime))
  
  observations_not_included_nback <- model_dat_nback %>% 
    anti_join(model_dat_filtered, by = c("id_esm", "day", "beep"))
  
  observations_not_included_nback %>%
  select(id_esm, day, beep, nback_dprime) %>%
  summarise(
    mean_nback_dprime   = mean(nback_dprime),
    median_nback_dprime = median(nback_dprime),
    sd_nback_dprime     = sd(nback_dprime),
    min_nback_dprime    = min(nback_dprime),
    max_nback_dprime    = max(nback_dprime)
  )
  
```


## statistical tests for 'not included' and 'final' samples

```{r}

# test for median age
  
  # normality test
  shapiro.test(dat_not_included$age)
  shapiro.test(model_dat_participants$age)
    # -> normality not assumed -> mann-whitney
  
  # mann-whitney test
  # wilcox.test(age ~ group, data = age_test)
  wilcox.test(dat_not_included$age, model_dat_participants$age)
  

# test for gender distribution
  gender_compare_not_included <- tibble(gender = dat_not_included$gender, group = "not included")
  gender_compare_included <- tibble(gender = model_dat_participants$gender, group = "included")
  gender_compare <- gender_compare_not_included %>% bind_rows(gender_compare_included)
  
  chisq.test(gender_compare$group, gender_compare$gender)


# test for education
  edu_compare_not_included <- tibble(edu = dat_not_included$edu, group = "not_included")
  edu_compare_included <- tibble(edu = model_dat_participants$edu, group = "included")
  edu_compare <- edu_compare_not_included %>% bind_rows(edu_compare_included)
  
  chisq.test(edu_compare$group, edu_compare$edu)
  

# test for go/no-go
   
  # normality test
  shapiro.test(model_dat_filtered$gonogo_dprime)
  shapiro.test(observations_not_included_gonogo$gonogo_dprime)
    # -> normality not assumed -> mann-whitney
  
  # mann-whitney test
  wilcox.test(model_dat_filtered$gonogo_dprime, observations_not_included_gonogo$gonogo_dprime)
  

# test for n-back
  
  # normality test
  shapiro.test(model_dat_filtered$nback_dprime)
  shapiro.test(observations_not_included_nback$nback_dprime)
    # -> normality not assumed -> mann-whitney

  # mann-whitney test
  wilcox.test(model_dat_filtered$nback_dprime, observations_not_included_gonogo$gonogo_dprime)
  
  
# test for event unpleasantness 
  
  # normality test
  shapiro.test(model_dat_filtered$event_unpleasantness)
  lillie.test(observations_not_included_ESM$event_unpleasantness)
    # -> normality not assumed -> mann-whitney

  # mann-whitney test
  wilcox.test(model_dat_filtered$event_unpleasantness, observations_not_included_ESM$event_unpleasantness)
  
  
```

# calculating compliance in unscreened sample

## compliance on ESM surveys

```{r}

# identify day of quitting

dat_unscreened_maxday_esm <- dat_cleaned %>% 
  select(id_esm, date_esm, day) %>% 
  filter(!is.na(date_esm)) %>% 
  group_by(id_esm) %>% 
  summarise(max_day_esm = max(day))

dat_cleaned <- dat_cleaned %>% 
  left_join(dat_unscreened_maxday_esm, by = "id_esm")


# compliance on ESM survey

dat_cleaned %>% 
  distinct() %>% 
  select(id_esm, date_esm, day, beep, max_day_esm) %>% 
  group_by(id_esm) %>%
  summarise(completed = sum(!is.na(date_esm)),
            all = max_day_esm * 8, 
            esm_compl_initial = (completed / all) * 100
  ) %>%
  ungroup() %>%
  summarise(median_initial_compl = median(esm_compl_initial),
            mean_initial_compl   = mean(esm_compl_initial),
            sd_initial_compl     = sd(esm_compl_initial),
            min_initial_compl    = min(esm_compl_initial),
            max_initial_compl    = max(esm_compl_initial))

```


## compliance on cogtasks

### go/no-go

```{r}

dat_cleaned %>%
  distinct() %>% 
  select(id_esm, start_date, day, beep, gonogo_dprime, max_day_esm) %>%
  group_by(id_esm) %>% 
  summarise(completed = sum(!is.na(gonogo_dprime)),
            all = max_day_esm * 8,
            gonogo_compl_initial = (completed / all) * 100
  ) %>%
  ungroup() %>%
  
  summarise(median_initial_compl = median(gonogo_compl_initial),
            mean_initial_compl   = mean(gonogo_compl_initial),
            sd_initial_compl     = sd(gonogo_compl_initial),
            min_initial_compl    = min(gonogo_compl_initial),
            max_initial_compl    = max(gonogo_compl_initial))

```


### n-back

```{r}

dat_cleaned %>%
  distinct() %>% 
  select(id_esm, start_date, day, beep, nback_dprime, max_day_esm) %>%
  group_by(id_esm) %>% 
  summarise(completed = sum(!is.na(nback_dprime)),
            all = max_day_esm * 8,
            nback_compl_initial = (completed / all) * 100
  ) %>%
  ungroup() %>%
  
  summarise(median_initial_compl = median(nback_compl_initial),
            meaninitiall_compl   = mean(nback_compl_initial),
            sd_initial_compl     = sd(nback_compl_initial),
            min_initial_compl    = min(nback_compl_initial),
            max_initial_compl    = max(nback_compl_initial))

```


# calculating compliance in final sample

## compliance on ESM surveys

```{r}

# filter participants who are included in model_dat, but keep all of their data, even rows not included in model_dat

dat_cleaned_model <- dat_cleaned %>% 
  semi_join(model_dat_filtered, by = "id_esm")


dat_cleaned_model %>%
  distinct() %>% 
  select(id_esm, date_esm, day, beep, max_day_esm) %>% 
  group_by(id_esm) %>%
  summarise(completed = sum(!is.na(date_esm)),
            all = max_day_esm * 8, 
            esm_compl_final = (completed / all) * 100
  ) %>%
  ungroup() %>%
  summarise(median_final_compl = median(esm_compl_final),
            mean_final_compl   = mean(esm_compl_final),
            sd_final_compl     = sd(esm_compl_final),
            min_final_compl    = min(esm_compl_final),
            max_final_compl    = max(esm_compl_final))

```


## compliance on cogtasks

### go/no-go

```{r}

dat_cleaned_model %>%
  distinct() %>% 
  select(id_esm, start_date, day, beep, gonogo_dprime, max_day_esm) %>%
  group_by(id_esm) %>% 
  summarise(completed = sum(!is.na(gonogo_dprime)),
            all = max_day_esm * 8,
            gonogo_compl_final = (completed / all) * 100
  ) %>%
  ungroup() %>%
 
  summarise(median_final_compl = median(gonogo_compl_final),
            mean_final_compl   = mean(gonogo_compl_final),
            sd_final_compl     = sd(gonogo_compl_final),
            min_final_compl    = min(gonogo_compl_final),
            max_final_compl    = max(gonogo_compl_final))

```


### n-back

```{r}
  
dat_cleaned_model %>%
  distinct() %>%
  select(id_esm, start_date, day, beep, nback_dprime, max_day_esm) %>%
  group_by(id_esm) %>%
  summarise(completed = sum(!is.na(nback_dprime)),
            all = max_day_esm * 8,
            nback_compl_final = (completed / all) * 100
  ) %>%
  ungroup() %>%
  
  summarise(median_final_compl = median(nback_compl_final),
            mean_final_compl   = mean(nback_compl_final),
            sd_final_compl     = sd(nback_compl_final),
            min_final_compl    = min(nback_compl_final),
            max_final_compl    = max(nback_compl_final))

```

