---
title: "R Notebook for posaff"
output: html_document
author: "Levente Rónai, Bertalan Polner"
editor_options: 
  markdown: 
    wrap: sentence
---

## Setup and load packages

```{r}

options(scipen = 999)

# Import and tidy data
library(tidyverse)
library(dplyr)
library(lme4)
library(lmerTest)
library(lavaan)
library(git2r)


# Modelling
# Ordinal regression
library(ordinal)
# Model visualisation, diagnostics and summary
library(sjPlot)
```

## Cumulative mixed models for positive affective states (cheerfulness and calmness)

Conducting the same model selection and #### Assumption checking process that one can see with the models predicting negative affective states

Models are pretty much okay with the assumptions; however issues occurred with complete separation, convergence and loglik/aic discrepancy in some cases, see the model fits below

```{r}
sessionInfo()
```

## Load and prepare data

```{r}
model_dat <- read.csv("processed_data/model_dat.csv")
```

## Convert data on emotional states into ordinal factors

```{r}
model_ord_dat <-
  model_dat %>% 
  mutate(
    across(c(angry, sad, cheerful, calm), .fns = ~factor(.x, levels = 1:6, ordered = TRUE))
  )
```

## Predicting cheerfulness

### Go/nogo

```{r}

# mod_clmm_cheerful_gonogo <-
#   model_ord_dat %>%
#   clmm(
#     cheerful ~ age + gender + beep + day +
#       event_unpleasantness_centered * gonogo_dprime_centered +
#       (gonogo_dprime_centered | id_esm ),
#     data = .
#   )
# 
# saveRDS(mod_clmm_cheerful_gonogo,"saved_models/ordinal_models/mod_clmm_cheerful_gonogo.rds")

### the model converges only with the random effect for gonogo scores

mod_clmm_cheerful_gonogo <- readRDS("saved_models/ordinal_models/mod_clmm_cheerful_gonogo.rds")

```

#### Assumption check

```{r}

# mod_clmm2_cheerful_gonogo <-
#   model_ord_dat %>% 
#   clmm2(
#     cheerful ~ 1 + event_unpleasantness_centered * gonogo_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# 
# saveRDS(mod_clmm2_cheerful_gonogo,"saved_models/ordinal_models/mod_clmm2_cheerful_gonogo.rds")
mod_clmm2_cheerful_gonogo <- readRDS("saved_models/ordinal_models/mod_clmm2_cheerful_gonogo.rds")

```

```{r}

# mod_clmm2_cheerful_gonogo_nominal <-
#   model_ord_dat %>% 
#   clmm2(
#     cheerful ~ 1,
#     nominal = ~ event_unpleasantness_centered * gonogo_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# 
# saveRDS(mod_clmm2_cheerful_gonogo_nominal,"saved_models/ordinal_models/mod_clmm2_cheerful_gonogo_nominal.rds")
mod_clmm2_cheerful_gonogo_nominal <- readRDS("saved_models/ordinal_models/mod_clmm2_cheerful_gonogo_nominal.rds")
```

#### AIC shows virtually no difference, while Likelihood-ratio test prefers the model with the nominal effects.

```{r}
anova(
  mod_clmm2_cheerful_gonogo,
  mod_clmm2_cheerful_gonogo_nominal
)

```

```{r}

AIC(
  mod_clmm2_cheerful_gonogo,
  mod_clmm2_cheerful_gonogo_nominal
)

```

### Nback

```{r}

# mod_clmm_cheerful_nback <-
#   model_ord_dat %>%
#   clmm(
#     cheerful ~ age + gender + # day + beep +
#       event_unpleasantness_centered * nback_dprime_centered +
#       (event_unpleasantness_centered | id_esm ),
#     data = .
#   )
# 
# saveRDS(mod_clmm_cheerful_nback,"saved_models/ordinal_models/mod_clmm_cheerful_nback.rds")

mod_clmm_cheerful_nback <- readRDS("saved_models/ordinal_models/mod_clmm_cheerful_nback.rds")

```

```{r}
tab_model(mod_clmm_cheerful_nback)
```

This model produces NaNs for confidence intervals and p values which is due to complete separation:

<https://stats.stackexchange.com/questions/231777/multinomial-logistic-regression-nans-produced-in-r-and-no-significant-variables>

<https://stats.oarc.ucla.edu/other/mult-pkg/faq/general/faqwhat-is-complete-or-quasi-complete-separation-in-logistic-regression-and-what-are-some-strategies-to-deal-with-the-issue/>

If day and beep excluded, the model converges

#### Assumption check

```{r}

# mod_clmm2_cheerful_nback <-
#   model_ord_dat %>% 
#   clmm2(
#     cheerful ~ 1 + event_unpleasantness_centered * nback_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# 
# saveRDS(mod_clmm2_cheerful_nback,"saved_models/ordinal_models/mod_clmm2_cheerful_nback.rds")
mod_clmm2_cheerful_nback <- readRDS("saved_models/ordinal_models/mod_clmm2_cheerful_nback.rds")
```

```{r}
# mod_clmm2_cheerful_nback_nominal <-
#   model_ord_dat %>% 
#   clmm2(
#     cheerful ~ 1,
#     nominal = ~ event_unpleasantness_centered * nback_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# 
# saveRDS(mod_clmm2_cheerful_nback_nominal,"saved_models/ordinal_models/mod_clmm2_cheerful_nback_nominal.rds")
mod_clmm2_cheerful_nback_nominal <- readRDS("saved_models/ordinal_models/mod_clmm2_cheerful_nback_nominal.rds")
```

```{r}
anova(
  mod_clmm2_cheerful_nback,
  mod_clmm2_cheerful_nback_nominal
)

```

```{r}

AIC(
  mod_clmm2_cheerful_nback,
  mod_clmm2_cheerful_nback_nominal
)

```

According to AIC, the nominal model is the better one whereas the loglik indicates the ordinal as the better fit: proportional odds assumption might be violated in this model

## Predicting calmness

### Go/nogo

```{r}

# mod_clmm_calm_gonogo <-
#   model_ord_dat %>%
#   clmm(
#     calm ~ age + gender + beep + day +
#       event_unpleasantness_centered * gonogo_dprime_centered +
#       (event_unpleasantness_centered * gonogo_dprime_centered | id_esm ),
#     data = .
#   )
# 
# saveRDS(mod_clmm_calm_gonogo,"saved_models/ordinal_models/mod_clmm_calm_gonogo.rds")

mod_clmm_calm_gonogo <- readRDS("saved_models/ordinal_models/mod_clmm_calm_gonogo.rds")

```

#### Assumption check

```{r}

# mod_clmm2_calm_gonogo <-
#   model_ord_dat %>% 
#   clmm2(
#     calm ~ 1 + event_unpleasantness_centered * gonogo_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# saveRDS(mod_clmm2_calm_gonogo,"saved_models/ordinal_models/mod_clmm2_calm_gonogo.rds")
mod_clmm2_calm_gonogo <- readRDS("saved_models/ordinal_models/mod_clmm2_calm_gonogo.rds")
```

```{r}
# mod_clmm2_calm_gonogo_nominal <-
#   model_ord_dat %>% 
#   clmm2(
#     calm ~ 1,
#     nominal = ~ event_unpleasantness_centered * gonogo_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# saveRDS(mod_clmm2_calm_gonogo_nominal,"saved_models/ordinal_models/mod_clmm2_calm_gonogo_nominal.rds")
mod_clmm2_calm_gonogo_nominal <- readRDS("saved_models/ordinal_models/mod_clmm2_calm_gonogo_nominal.rds")
```

```{r}
anova(
  mod_clmm2_calm_gonogo,
  mod_clmm2_calm_gonogo_nominal
)

```

```{r}

AIC(
  mod_clmm2_calm_gonogo,
  mod_clmm2_calm_gonogo_nominal
)

```

### Nback

```{r}

# mod_clmm_calm_nback <-
#   model_ord_dat %>%
#   clmm(
#     calm ~ age + gender + beep + day +
#       event_unpleasantness_centered * nback_dprime_centered +
#       (event_unpleasantness_centered * nback_dprime_centered | id_esm ),
#     data = .
#   )
# 
# saveRDS(mod_clmm_calm_nback,"saved_models/ordinal_models/mod_clmm_calm_nback.rds")

mod_clmm_calm_nback <- readRDS("saved_models/ordinal_models/mod_clmm_calm_nback.rds")


```

#### Assumption check

```{r}

# mod_clmm2_calm_nback <-
#   model_ord_dat %>% 
#   clmm2(
#     calm ~ 1 + event_unpleasantness_centered * nback_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# saveRDS(mod_clmm2_calm_nback,"saved_models/ordinal_models/mod_clmm2_calm_nback.rds")
mod_clmm2_calm_nback <- readRDS("saved_models/ordinal_models/mod_clmm2_calm_nback.rds")
```

```{r}
# mod_clmm2_calm_nback_nominal <-
#   model_ord_dat %>% 
#   clmm2(
#     calm ~ 1,
#     nominal = ~ event_unpleasantness_centered * nback_dprime_centered,
#     random =  factor(id_esm), 
#     data = .,
#     Hess = TRUE
#   )
# saveRDS(mod_clmm2_calm_nback_nominal,"saved_models/ordinal_models/mod_clmm2_calm_nback_nominal.rds")
mod_clmm2_calm_nback_nominal <- readRDS("saved_models/ordinal_models/mod_clmm2_calm_nback_nominal.rds")
```

```{r}
anova(
  mod_clmm2_calm_nback,
  mod_clmm2_calm_nback_nominal
)
```

```{r}

AIC(
  mod_clmm2_calm_nback,
  mod_clmm2_calm_nback_nominal
)

```

```{r}
tab_model(mod_clmm_cheerful_gonogo, mod_clmm_cheerful_nback, mod_clmm_calm_gonogo, mod_clmm_calm_nback)
```
