---
title: "Within-session reliability of cognitive control tasks"
output: html_document
author: "Levente Rónai, Bertalan Polner"

---

```{r}
library(tidyverse)
library(psych)
library(lme4)
library(sjPlot)
```

# between-individual reliability

```{r}

model_dat <- 
  read.csv("processed_data/model_dat.csv")

```

```{r}

icc_gng <- lmer(gonogo_dprime ~ 1 + (1 | id_esm), data = model_dat)
icc_nb  <- lmer(nback_dprime ~ 1 + (1 | id_esm), data = model_dat)


tab_model(
  icc_gng,
  icc_nb,
  
  show.ngroups = TRUE,  
  p.val = "satterthwaite", 
  show.aic = TRUE,
  show.aicc = TRUE,
  p.style = "stars",
  digits = 2, 
  digits.re = 3, 
  emph.p = TRUE,  
  string.ci = "95% CI",
  dv.labels = 
    c("gonogo_dprime",
      "nback_dprime")
)

```

# Prepare data

```{r}
dat_cogtask <- 
  read_csv("processed_data/dat_cogtask_raw.csv")

```

```{r}
filtered <-
  read_csv("processed_data/dat_cogtask.csv")
```
Keep raw cogtask data only for those sessions that survived screening 

```{r}
dat_cogtask <- 
  dat_cogtask %>% 
  semi_join(filtered, 
            by = 
              c("session_esm" = "session_esm", 
                "date_cogtask" = "date_cogtask_orig")
            ) 
```



# Calculate split-half reliability

## Function to calculate key task metrics

```{r}
calc_task_indices <- function(dat){
  
  dat <-
    dat %>%
    mutate(
      corr_rt = ifelse(trial_type == 1, rt, NA)
    ) %>%
    group_by(session_esm, date_cogtask) %>% 
    summarise(
      n_of_trials    = mean(count_of_trial),
      task           = max(as.character(task)),
      corr_rt_median = median(corr_rt, na.rm = T),
      n_hit          = sum(trial_type == 1 & !is.na(rt)),
      n_fa           = sum(trial_type == 0 & !is.na(rt)),
      n_miss         = sum(trial_type == 1 & is.na(rt)),
      n_cr           = sum(trial_type == 0 & is.na(rt)),
      n_targets      = n_hit + n_miss,
      n_distractors  = n_fa + n_cr
    ) %>% 
    ungroup()
  
  dprime <- 
    psycho::dprime(
      n_hit         = dat$n_hit,
      n_fa          = dat$n_fa,
      n_miss        = dat$n_miss,
      n_cr          = dat$n_cr,
      n_targets     = dat$n_targets,
      n_distractors = dat$n_distractors)
  
  
  dat <-
    dat %>%
    cbind(dprime) %>% 
    select(everything(), -c(n_hit:n_distractors), -c(aprime:c))   %>%
    group_by(session_esm, task) %>%
    mutate_at(vars(corr_rt_median, dprime, beta),
              ~scale(., scale = FALSE)
    ) %>%
    ungroup()
  
  return(dat)
}

```

## Function to calculate split-half reliability using Spearman-Brown formula

```{r}
split_half_sb <- function(dat){
  
  metrics <- c("dprime", "beta", "corr_rt_median")
  
  task_name <- dat$task
  
  # Initilize vector for results
  split_half_corr <- numeric()
  
  # Calculate correlation between metrics computed in the two halves
  for(metric in metrics){
    split_half_corr[str_c(task_name, "_", metric )] <- 
      dat %>% 
      select(contains(metric)) %>% 
      cor(use = "pairwise.complete.obs") %>% 
      .[1,2]
  }
  # Apply Spearman-Brown formula
  split_half_reliability <- split_half_corr*2/(1 + split_half_corr)  
  
  
}
```

```{r, warning=FALSE, message=FALSE}
# n_samples <- 100
# res <- list()
# for(i in 1:n_samples){
# 
# 
#   # Create random halves
# 
#   slices <- list()
# 
#   slices[[1]] <-
#     dat_cogtask %>%
#     group_by(session_esm, task, date_cogtask, trial_type) %>%
#     slice_sample(prop = 0.5) %>%
#     ungroup()
# 
#   slices[[2]] <-
#     dat_cogtask %>%
#     anti_join(slices[[1]])
# 
#   # Calculate task indices in both halves
# 
#   dat_splithalves <-
#     map(slices, calc_task_indices) %>%
#     reduce(left_join, by = c("session_esm", "date_cogtask", "task"))
# 
# 
#   # Compute split-half reliability and store the result
# 
#   res[[i]] <-
#     dat_splithalves %>%
#     group_split(task) %>%
#     map(split_half_sb) %>%
#     unlist()
# 
#   print(str_c("Iteration ", i, " done"))
# 
# }
# 
# res <-
#   res %>%
#   reduce(bind_rows)
# 
# write_rds(res, "processed_data/cogtask_reliablity_splithalf_centered.RDS")
```

Results after centering

```{r}
read_rds("processed_data/cogtask_reliablity_splithalf_centered.RDS") %>% 
  summary(res)
```

For comparison: reliability of non-centered cogtask scores

```{r}
read_rds("processed_data/cogtask_reliablity_splithalf_notcentered.RDS") %>% 
  summary()
```

# Dataviz

```{r}
slices <- list()
slices[[1]] <- 
  dat_cogtask %>% 
  group_by(session_esm, task, date_cogtask, trial_type) %>% 
  slice_sample(prop = 0.5) %>% 
  ungroup()
slices[[2]] <- 
  dat_cogtask %>% 
  anti_join(slices[[1]])
# Calculate task indices in both halves
dat_splithalves <- 
  map(slices, calc_task_indices) %>% 
  reduce(left_join, by = c("session_esm", "date_cogtask", "task")) 
```

```{r}
dat_splithalves %>% 
  group_by(session_esm, task) %>% 
  summarise(
    cor_splithalf = cor(dprime.x, dprime.y),
    n_sessions    = n()
  ) %>% 
  ggplot(aes(cor_splithalf)) + 
  geom_histogram() +
  facet_grid( task ~ . ) + 
  xlab("Correlation between two random halves of the same session' dprime scores")
```


```{r, fig.width=15, fig.height=15}
dat_splithalves %>%
  filter(task == "go-nogo") %>% 
  ggplot(aes(dprime.x, dprime.y)) + 
  geom_jitter() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap( ~session_esm)+ 
  theme(
    strip.background = element_blank(),
    strip.text.x     = element_blank(),
    plot.title       = element_text(size = 20, hjust = 0.5)
  ) + 
  ggtitle("Within-person reliability of go-nogo d\'") + 
  xlab("d\' for random half of a session") + ylab("d\' for the other half of the session")
```

```{r, fig.width=15, fig.height=15}
dat_splithalves %>%
  filter(task == "n-back-number") %>% 
  ggplot(aes(dprime.x, dprime.y)) + 
  geom_jitter() + 
  geom_smooth(method = "lm", se = FALSE, color = "purple") + 
  facet_wrap( ~session_esm)+ 
  theme(
    strip.background = element_blank(),
    strip.text.x     = element_blank(),
    plot.title       = element_text(size = 20, hjust = 0.5)
  ) + 
  ggtitle("Within-person reliability of n-back d\'") + 
  xlab("d\' for random half of a session") + ylab("d\' for the other half of the session")
```
